---
title: "P8130_HW3_Problem4"
author: "Chen Shaoxuan"
date: "11/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r results = "hide", message = FALSE, echo = FALSE}
library(readr)
library(tidyverse)
library(janitor)
library(ggplot2)
library(ggthemes)
library(ggridges)
library(haven)
library(knitr)
library(forcats)
```

```{r results = "hide", message = FALSE, echo = FALSE}
who_life_exp_data = read_csv("WHO_LifeExpectancy.csv") %>% 
  clean_names()
```

##a).Describe dataset and the variables of interest.
This dataset is the Global Health Observatory Data in 2012, which contains 137 entries and 12 variables of interest. The variables are country, pcgovhealth(per capital government health expend), legislation(IHR Legislation Score), coordination(IHR Coordination Score), surveillance(IHR Surveillance Score), preparedness(IHR Preparedness Score), foodsafety(IHR Food Safety Score), birthrate(Birth Rate per 1000), mlifebirth(Male Life Expectancy at ), flifebirth(Female Life Expectancy at), lifebirth(Life Expectancy at Birth), life60(Life Expectancy at Age 60). We are interested in the female life expectancy and the government health expend.

##b).Summary statistics for each of the variables of interest
```{r echo = FALSE}
who_life_exp_data %>% 
  select(pcgovhealth, flifebirth) %>% 
  gather(key = "variables", value = "number", pcgovhealth:flifebirth) %>% 
  group_by(variables) %>% 
  summarise(n_obs = n(),
             mean = round(mean(number), digits = 2),
               sd = round(sd(number), digits = 2),
           median = median(number),
          munimum = min(number),
           maxiun = max(number),
        first_Qu. = quantile(number, 0.25),
        third_Qu. = quantile(number, 0.75)
           ) %>% 
  knitr::kable()
```

##c) 
###Graphical displays (histograms, boxplots, and scatterplots)
```{r echo = FALSE}
plot_data = who_life_exp_data %>% 
  select(pcgovhealth, flifebirth) %>% 
  gather(key = "variables", value = "number", pcgovhealth:flifebirth) %>% 
  mutate(variables = as.factor(variables)) %>%
  group_by(variables) %>% 
  summarise(sum = sum(number)) %>% 
  mutate(variables = fct_reorder(variables, sum))
#Histogram
who_life_exp_data %>%
ggplot(aes(x = pcgovhealth)) +
geom_histogram(binwidth = 67, color = "blue") + xlab("per capital government health expend")  
  
who_life_exp_data %>%
ggplot(aes(x = flifebirth)) +
geom_histogram(binwidth = 0.5, color = "blue") + xlab("Female life expectancy at birth (year)")

#Boxplots
who_life_exp_data %>%
ggplot(aes(x ="government health expend" ,y = pcgovhealth)) +
geom_boxplot(color = "blue") + xlab("per capital government health expend distribution") 

who_life_exp_data %>%
ggplot(aes(x ="Female life expectancy " ,y = flifebirth)) +
geom_boxplot(color = "blue") + xlab("Female life expectancy distribution") 

#Scatterplot
who_life_exp_data %>%
ggplot(aes(x = pcgovhealth ,y = flifebirth)) +
geom_point(color = "blue") + xlab("Distribution of per capital govern. health expend and female life expectancy") 

```

###i) Is linear regression a reasonable model for this problem? Explain why or why not.  
   *From the plots, it is obvious that linear regression is not a reasonable model for this problem. Since with the increase of the capital of government expend, the female life expectancy doesn't have a linear increase, i.e., steady growth. Form the scatterplot, at first, the female life expectancy increase very quick with the increase of government health expend. However, after 2000, female life expectancy increase very slowly.
   
   
###ii) Try a natural logarithm transformation of the predictor and check the relationship.
```{r}
log_data = who_life_exp_data %>% 
  na.omit() %>% 
  gather(key = "variables", value = "number", pcgovhealth:life60) %>%
  mutate(number = number + 1 ) %>% 
  mutate(number = log(number)) %>% 
  mutate(variables = as.factor(variables)) %>% 
  group_by(variables) %>% 
  summarise(sum = sum(number)) %>% 
  mutate(variables = fct_reorder(variables, sum)) 

log_data %>% 
 ggplot(aes(x = variables, y = sum, fill = variables)) +
  geom_col() +
   xlab("variables") + ylab("number") +
   ggtitle("Histogram of the distribution of each variable of interest") +
  theme(legend.position = "bottom" ,
        plot.title = element_text(hjust = 0.5),              ######### make the title in the middle 
        axis.text.x = element_text(angle = 90, hjust = 0.9))
```

    *This time, we can see the obvious linear relationships from between different variables form the plot.

##d). ) Fit a linear regression where female life expectancy in a country is regressed on per capita spending on health (original or transformed). (8p) 

Note: If you chose a transformation of the predictor, make sure you consider it in all aspects, including interpretation. 

```{r}
#At first check the relationship between "pcgovhealth" and "flifebirth"
reg_lm1 = lm(data = who_life_exp_data,flifebirth~pcgovhealth)
plot(who_life_exp_data$flifebirth, who_life_exp_data$pcgovhealth)
abline(reg_lm1, lwd = 2, col = 2)
```

It is obvious that female life expectancy is not regressed on per capita spending on health. Thus, I choose to do logarithm transformation of the predictor.

```{r}
pcgovhealth_flifebirth_lm_data = who_life_exp_data %>% 
  select(pcgovhealth, flifebirth) %>%
  mutate(pcgovhealth = log(pcgovhealth)) 
##Check
reg_lm = lm(data = pcgovhealth_flifebirth_lm_data, flifebirth~pcgovhealth)
plot(pcgovhealth_flifebirth_lm_data, flifebirth~pcgovhealth)
abline(reg_lm, lwd = 2, col = 2)
 #pcgovhealth_flifebirth_lm_data %>% 
  #ggplot(aes(x = pcgovhealth, y = flifebirth, color = pcgovhealth)) +
  #geom_point(alpha = .8) +
   #ggtitle("Plot after logarithm transformation of the predictor ") +
  #theme(legend.position = "bottom" ,
   #     plot.title = element_text(hjust = 0.5))
```

####fit the regression model
```{r}
reg_lm = lm(data = pcgovhealth_flifebirth_lm_data, flifebirth~pcgovhealth)
summary(reg_lm)
```

###i) Report regression results including model coefficients, standard errors, confidence intervals (for the model parameters), test statistics and the associated p-values.

```{r}
#Calculate the 95% CI for moder parameters?
##Beta1
CI_data = pcgovhealth_flifebirth_lm_data 
colnames(CI_data) <- c("x", "y")
x = CI_data$x
y = CI_data$y
b_1 = sum((x-mean(x))*(y-mean(y))) / (sum((x-mean(x))^2))
b_0 = mean(y) - b_1*mean(x)
SSE = sum((y - b_0 - b_1*x)^2)
MSE = SSE / 135
s_b1 = sqrt(MSE/(sum((x-mean(x))^2)))
s_b0 = MSE*(1/137 + (mean(x)^2)/(sum((x-mean(x))^2)))

###95% CI for b_1
CI_b_1_upper_limit = b_1 + qt(0.95, 135)*s_b1
CI_b_1_lower_limit = b_1 - qt(0.95, 135)*s_b1
###95% CI for b_0
CI_b_0_upper_limit = b_0 + qt(0.95, 135)*s_b0
CI_b_0_lower_limit = b_0 - qt(0.95, 135)*s_b0

```

  As for the regression results, the intercept is 55.3524, which represents$\beta_0$ in regression model. The second row in the coefficient is slope(3.5887), which represents $\beta_1$. Standard error is 0.2372, t-value is 15.13, P-value is 2e^-16. The The 95% CI for $\beta_1$ is (3.196, 3.589). The 95% CI for  $\beta_0$ is (52.338, 58.337)

### ii) Also, write the ‘best-fitting’ line equation. 
The 'best-fitting' line equation is $\hat{Y_i}= 3.5887ln\hat{X_i} + 55.3524$

###iii) Comment on your findings and interpret the estimated slope of the model. 
I find $\hat{\beta_1}$ is very close to the upper limit of 95% CI of $\beta_1$. 
The interpretation of the estimate slope is that with every e (about2.718) increase capita spending on health, the female life expectancy will increase 3.5887 years.

###e) Provide a point estimate and a 95% interval for the mean life expectancy for countries that have per capita spending on health of approximately 1,100 US dollars. Is this a confidence or a prediction interval? Make sure you provide its interpretation correctly. (3p) 

```{r}
plot(who_life_exp_data$flifebirth, who_life_exp_data$pcgovhealth)
abline(reg_lm1, lwd = 2, col = 2)

#data_hosp = as.data.frame(data_hosp)
reg_lm = lm(data = pcgovhealth_flifebirth_lm_data, flifebirth~pcgovhealth)
#create a new data frame that set the beds value.
newdata=data.frame(pcgovhealth = log(1100))  
#Apply the predict function and set the predictor variable in the newdata argument. We also set the interval type as "predict", and use the default 0.95 confidence level.
predict(reg_lm, newdata, interval ="prediction", level = 0.95) 
```

It is a prediction inteval. With per capita spending on health of approximately 1,100 US dollars，we have 95% confidence that the 

